### 2024-11-30 background.tex
init commit

\chapter{Background and Related Work}
\label{background}

Background 123

% \begin{figure*}
% \centering
% \includegraphics[width=0.98\textwidth]{Chapter1/Images/Fig1.pdf}
% \caption{caption}
% \label{figCh1:figure}
% \end{figure*}

\renewcommand\thefigure{\thechapter .\arabic{figure}}