### 2024-11-30 conclusions.tex

\chapter{Conclusion}
\label{concl}

asdfsdfasdf
\newpage
asdasasfdfdf

% \begin{figure*}
% \centering
% \includegraphics[width=0.98\textwidth]{Chapter1/Images/Fig1.pdf}
% \caption{caption}
% \label{figCh1:figure}
% \end{figure*}

\renewcommand\thefigure{\thechapter .\arabic{figure}}