### 2024-11-30 preliminaries.tex

\chapter{Preliminaries}
\label{prelim}

Preliminaries 123

\renewcommand\thefigure{\thechapter .\arabic{figure}}