### 2025-05-19 introduction.tex

\chapter[Introduction]{Introduction}
\label{chap:1}

The idea of a formalised mechanist is simple but powerful: to trust mathematics in software, we must mechanise not only calculations but also proofs. A mechanist builds a logic engine where every theorem arises from a small set of primitive rules. If those rules are sound, everything derived from them inherits that soundness.

HOL4 supports this mechanised philosophy \cite{slind2008brief}. It encodes logic using a typed \(\lambda\)-calculus and restricts theorem creation to a protected MetaLanguage (ML) type, ensuring that every step is checked with mathematical rigour.

In this thesis, we formalise one of the most important results in probability theory: the Central Limit Theorem (CLT). When many independent effects accumulate, their combined influence often resembles a bell curve. The CLT explains this universal phenomenon observed across diverse fields—from biology and economics to physics and computation.

We formalise the Lyapunov version of the CLT from \cite{Chung:2001}, using Lindeberg’s replacement method. Let $\{X_j\}_{j=1}^n$ be a sequence of independent random variables with zero mean, finite variance, and third absolute moment:

\[
\mathbb{E}[X_j] = 0,\quad \mathbb{E}[X_j^2] = \sigma_j^2,\quad \mathbb{E}[|X_j|^3] = \gamma_j < \infty.
\]

Define the sum \( S_n = \sum_{j=1}^n X_j \), total variance \( s_n^2 = \sum \sigma_j^2 \), and total third moment \( \Gamma_n = \sum \gamma_j \). Then, under the Lyapunov condition:

\[
\lim_{n \to \infty} \frac{\Gamma_n}{s_n^3} = 0,
\]

we have convergence in distribution:
\[
\frac{S_n}{s_n} \xrightarrow{d} \mathcal{N}(0, 1).
\]

Here is the Lyapunov's form of the Central Limit Theorem in HOL4:
\begin{hol}
  \begin{alltt}
    Theorem central\_limit\_theorem :
    \(\!\!\!{\turn}\!\!\!\!\) \(\forall\)p X N.
    prob\_space p \(\land\)
    ext\_normal\_rv N p 0 1 \(\land\)
    (\(\forall\)i. real\_random\_variable (X i) p) \(\land\)
    (\(\forall\)n. indep\_vars p X (\(\lambda\)i. Borel) (count n)) \(\land\)
    (\(\forall\)i. expectation p (X i) = 0) \(\land\)
    (\(\forall\)i. expectation p (\(\lambda\)x. (abs (X i x))\(^3\)) < \(+\infty\)) \(\land\)
    (\(\forall\)i. variance p (X i) < PosInf) \(\land\)
    (\(\forall\)i. variance p (X i) \(\ne\) 0) \(\land\)
    (\(\forall\)n. (sqrt (second\_moments p X n)) \(\ne\) 0) \(\land\)
    ((\(\lambda\)n. (third\_moments p X n) / ((sqrt (second\_moments p X n)) pow 3)) --> 0) sequentially
    \(\Rightarrow\)
    ((\(\lambda\)n x. (SIGMA (\(\lambda\)i. X i x) (count n)) / (sqrt (second\_moments p X n))) --> N)
    (in\_distribution p)
  \end{alltt}
\end{hol}

Although the final convergence proof is not yet complete, this thesis represents a significant step toward a full mechanisation. Over 6000 lines of HOL4 code were developed, including formal definitions, theorems, and supporting libraries. An initial attempt using moment-generating functions was abandoned due to insufficient HOL4 support. The current Taylor-based approach has allowed us to formalise all major components. The only missing piece is a bound involving the gamma function, which requires integration techniques not yet supported in HOL4. This is a matter of technical realisation, not conceptual difficulty.