### 2025-05-22 abstract.tex
Minor changes

\chapter*{Abstract}
\label{abstract}

This thesis presents a formalisation of Lyapunov’s Central Limit Theorem (CLT) in the HOL4 theorem prover. The proof follows the Lindeberg replacement method and employs Taylor expansion techniques to quantify the difference between the distribution of normalised partial sums and the standard normal distribution.

The formalisation enriches HOL4’s existing \texttt{ProbabilityTheory} by introducing additional infrastructure necessary for the proof, including higher-order derivatives, bounds on smooth functions, and properties of moment conditions.

A key outcome of this work is a modular and rigorous development of the Lindeberg-based proof of the CLT, which complements existing formal efforts in \texttt{ProbabilityTheory} in HOL4. The approach demonstrates that classical analytical techniques can be accurately mechanised using modern theorem proving tools.