### 2025-05-22 acknowledgments.tex
Minor changes

\chapter*{Acknowledgments}
\label{ack}

The author want to give special thanks to Dr. \emph{Chun Tian} for his supervision of this thesis and for suggesting the topic of formalizing the Central Limit Theorem. Without Dr. Tian’s generous support, deep insights, and continuous guidance, the author wouldn't have been able to pursue this work within the HOL4 theorem prover and complete the requirements for graduation.

The author also wants to thanks to her family for their unconditional love, belief in her, and emotional support, even from afar. Their presence—whether near or distant—gave her strength through every challenge.

To her dear friend \emph{Le Manh Tai}, thank you for always being there—for every late night, every deadline, and every quiet moment of doubt. Your support made a difference more than words can say.

Finally, the author would like to thank all her friends in Vietnam, Australia, and around the world who have supported her in different ways. Without all of you, the author wouldn’t have made it this far.